---
title: "Lab 2 Lydia"
author: "Lydia Bleifuss"
date: "1/23/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE) #this is global code! You can set code chunk stuff for all up here (not messages or warning unless ou override it later on)


library(tidyverse)
library(janitor)
library(here)
library(here)
library(tmap)
library(sf)

```


##Binary logistic regression: Penguins! 

```{r}
gentoo <- read_csv("gentoo_lter.csv")
chinstrap <- read_csv("chinstrap_lter.csv")

#these dataframes have the same column headings, when in doubt, the most complete way to jain dataframes is "fulljoin", can always use drop NA etc etc later

penguins <- full_join(chinstrap, gentoo) %>% #basically put in tidy format, just added one dataframe on top of the other
  clean_names() %>% 
  mutate(sex = str_to_lower(sex)) %>% 
#usually used to add new columns on a dataframe, but can also overright an existing column with a change that we specify, which is what is done here (using the same column name "sex" to overright rather than "Sex" or something like that)
  filter(sex %in% c("male", "female")) #we want to only KEEP rows that say male or femal in the sex column and exclude anything else. The %in% opperator is a more elegant "or" statement, "look for anything in sex column that matches anything in this string, other way to do this is: filter (sex == "male" | sex = "female))

```


##Now, what trends do we see/initial impression (let's graph it!)

```{r}
ggplot(data = penguins, aes(x = sex, y = body_mass_g)) +
  geom_jitter(
    aes(color = sex, 
        pch = sex)
  ) +
  facet_wrap(~species)
```


##Code Chinstrap penguins as (1) and gentoo penguins as (0):

```{r}
penguin_bin <- penguins %>% 
  mutate(
    sp_binary = case_when(#creating something that is conditional on something else (this is where we are using if-then statements)
    species == "Chinstrap penguin (Pygoscelis antarctica)" ~ 1,
    species == "Gentoo penguin (Pygoscelis papua)" ~ 0)
    #T ~ "cool") # (~) means as a function of, this is an if-then statment, "T ~ "cool" which means if anything else is true/there, then label it cool (we don't have anything else but included for notes)
  )
```



#We are going to find the probabilities of a penguin being a (1) or a Chinstrap

Now Let's run our binary logistic regression 

```{r}
penguin_blr <- glm(sp_binary ~ sex + body_mass_g,
                   family = binomial,
                   data = penguin_bin)

summary(penguin_blr)
```

###Let's  make some predictions for penguins (classification...kinda):

1. What is the probability that a penguin is a Chinstrap penguin, if it weighs 4500 grams and is male? (based on graph, thinking it will likley be a chinstrap)

```{r}
df_m4500 <- data.frame(sex = "male", body_mass_g = 4500 ) #creating a new df with two columns, name of columns have to match what the model was given (varible name and entires have to match what the model was given) (= is assigning, and == is asking if this matches that and then true or false back) #numbers don't need quotations around them because they are numbers and not character strings

# FInd log odds of this penguin being Chinstrap: 

m4500_logodds <- predict(penguin_blr, newdata = df_m4500, type = "link") #predict funtion works for several types of models | "link" is default for glm is log odds

m4500_logodds #4.047 

m4500_prob <- predict(penguin_blr, newdata = df_m4500, type = "response") #this creates log odds to probability which is easier to understand

m4500_prob #1, 98%, therefore if we foudn a male on the isalnd that was 4500 grams, then there is a 98% probability that it is a chin strap
```


2. What is the probability that a penguin is a Chinstrap, if it is recorded as femal of 4000 g? (based on graph, 50/50?)

```{r}
df_f4000 <- data.frame(sex = "female", body_mass_g = 4000)

f4000_prob <- predict(penguin_blr, newdata = df_f4000, type = "response") 

f4000_prob #1, 65%, so, there is a 65% chance that a female found on the island that is 4000g is a chinstrap 
```

##How robust are these probabilities? Let's check on uncertainty...

Now let's make an entire data frame & use it to make then visualize logistic regression outcomes: 

```{r}
penguins_mock <- data.frame(
  body_mass_g = rep(seq(3000, 6000, length = 200), 2), #length is 200 rows or values ranging from 3000 - 6000, a sequence, you could say by = # rather than length if you prefer that, and we want to repeat this sequence (rep) 2 times which means penguins_mock will now have 400 observations
  sex = c(rep("male", 200), rep("female", 200))) #now, in this sequence, there are 200 female observations and 200 male observations

```

Now let's use those to make predictions for all of them:

```{r}
full_predict <- predict(penguin_blr, newdata = penguins_mock, type = "response", se.fit = TRUE) #se.fit=TRUE has the predicted fitted probabilities and the standard error, we are going to take those can get them with the other statistics, GET THOSE INTO A DATA FRAME , ALONG WITH THE MOCK DATA USED TO CREATE THEM:

final_df <- data.frame(
  penguins_mock,
  full_predict$fit,
  full_predict$se.fit
)

colnames(final_df) <- c("penguin_mass", "sex", "probability", "standard_error")
```


NOw let's graph our mock data and predicted probabilities!

```{r}
ggplot(data = final_df, aes(x = penguin_mass, y = probability)) +
  geom_line(aes(color = sex))

#shows the prob of a penguin being a chinstrap for the different masses and whether they are male or female 
#shift from a penguin being more likely a gentoo and chinstrap happens around 4000 for female and 4800 for male 

ggplot(data = final_df, aes(x = penguin_mass, y = probability)) +
  geom_line(aes(color = sex)) + 
  geom_ribbon(aes(ymin = probability - standard_error,
                  ymax = probability + standard_error, 
                  fill = sex),
                  alpha = 0.3) #STANDARD ERROR BAR!
```


##NOW some spatial data with sf

Read in teh spatial data using read_sf():

```{r}
cougars <- read_sf(dsn = here("cougar_connections"), layer = "ds1014") %>%  #spatial information is in the last column, the geometry column (has all of the coordinates that make up a polygon) in the sf package, the coordinates are sticky 
  clean_names()

large_corridors <- cougars %>% 
  filter(area_ac > 20000) %>% 
  select(area_ac) #only keep that (area_ac) column, but what really happens is that area_ac exists BUT geometry (SPACIAL) column is stuck to it 

plot(large_corridors) #you really don't want to create polygons in r, but can definitely use lat/long data 
```


```{r}
cougar_sub <- cougars %>%
  select(elev_mean) #mean elevation for all cooridors
```


#Can work in ggplot to creat static maps

```{r}
ggplot(data = cougar_sub) +
  geom_sf(aes(fill = elev_mean),
          color = NA)  ###NOW let's add a base layer and make it interactive!
```


```{r}
#Tmap can default make static map or view which will set it to interactive mapping

tmap_mode("view")

tm_shape(cougar_sub) +  #tells tmap what data you are using
 tm_fill("elev_mean") #leaf extras to change BASE MAPS??
```

